---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

# Source: https://stackoverflow.com/questions/34983822/how-to-have-r-formattable-rendered-to-pdf-output-and-how-to-have-percents-in-the
export_formattable <- function(f, file, width = "100%", height = NULL, 
                               background = "white", delay = 0.2)
{
  w <- as.htmlwidget(f, width = width, height = height)
  path <- htmltools::html_print(w, background = background, viewer = NULL)
  url <- paste0("file:///", gsub("\\\\", "/", normalizePath(path)))
  webshot::webshot(url,
          file = file,
          selector = ".formattable_widget",
          delay = delay)
}
```

# Installing the Package

You can install `melig` from Github with:

```R
# install.packages("remotes")

# Install the development version
remotes::install_github("jjchern/melig")

# Install the released version
remotes::install_github("jjchern/melig@v0.1.0")

# To uninstall the package, use:
# remove.packages("melig")
```

# About `melig`

[![Travis-CI Build Status](https://travis-ci.org/jjchern/melig.svg?branch=master)](https://travis-ci.org/jjchern/melig)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/jjchern/melig?branch=master&svg=true)](https://ci.appveyor.com/project/jjchern/melig)

The R package `melig` contains information on **M**edicaid income **elig**ibility limits, collected by [the Kaiser Family Foundation](http://kff.org/data-collection/trends-in-medicaid-income-eligibility-limits/). It includes the following datasets:

- `children.rda`: 
    - [Medicaid Income Eligibility Limits for Infants Ages 0–1, 2000–2023](https://www.kff.org/medicaid/state-indicator/medicaid-income-eligibility-limits-for-infants-ages-0-1/)
    - [Medicaid Income Eligibility Limits for Children Ages 1–5, 2000–2023](https://www.kff.org/medicaid/state-indicator/medicaid-income-eligibility-limits-for-children-ages-1-5/)
    - [Medicaid Income Eligibility Limits for Children Ages 6–18, 2000–2023](https://www.kff.org/medicaid/state-indicator/medicaid-income-eligibility-limits-for-children-ages-6-18/)
    - [Separate Children’s Health Insurance Program (CHIP) Income Eligibility Limits for Children, 2000–2023](https://www.kff.org/medicaid/state-indicator/separate-childrens-health-insurance-program-chip-income-eligibility-limits-for-children/)
    - [Medicaid/CHIP Upper Income Eligibility Limits for Children, 2000–2023](https://www.kff.org/medicaid/state-indicator/medicaidchip-upper-income-eligibility-limits-for-children/)
    
- `prenant_wom.rda`: [Medicaid and CHIP Income Eligibility Limits for Pregnant Women, 2003–2023](https://www.kff.org/medicaid/state-indicator/medicaid-and-chip-income-eligibility-limits-for-pregnant-women/)
- `parents.rda`: [Medicaid Income Eligibility Limits for Parents, 2002–2023](https://www.kff.org/medicaid/state-indicator/medicaid-income-eligibility-limits-for-parents/)
- `childless_adults.rda`:  [Medicaid Income Eligibility Limits for Other Non-Disabled Adults, 2011–2018](https://www.kff.org/medicaid/state-indicator/medicaid-income-eligibility-limits-for-other-non-disabled-adults/)

The datasets are stored in tidy format (long table).

The [`/data-raw/`](https://github.com/jjchern/melig/tree/master/data-raw) folder contains raw data and R scripts.

# Useful Links

- [Medicaid Income Eligibility Limits (Most Recent Data)](http://kff.org/state-category/medicaid-chip/medicaidchip-eligibility-limits/)
- [Annual Updates on Eligibility Rules, Enrollment and Renewal Procedures, and Cost-Sharing Practices in Medicaid and CHIP](http://kff.org/medicaid/report/annual-updates-on-eligibility-rules-enrollment-and/)

# Usage

## Long and Wide Formats

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(formattable)

# the datasets have been tidied
melig::parents

# but you can covert it back to the original wide format
melig::parents %>% 
  unite_("monyear", c("month", "year"), sep = " ") %>% 
  spread(monyear, cutoff)
```

## States that Have Income Cutoffs Greater Than 100 Federal Poverty Guidelines in 2014

```{r, message=FALSE, warning=FALSE}
# for parents
melig::parents %>% 
  filter(year == 2014) %>% 
  filter(cutoff >= 100) %>% 
  select(state, fips, usps, pa_cutoff = cutoff) %>% 
  formattable(list(pa_cutoff = normalize_bar("lightgreen", 0.2)))

# for childless adults
melig::childless_adults %>% 
  filter(year == 2014) %>% 
  filter(cutoff >= 100) %>% 
  select(state, fips, usps, ca_cutoff = cutoff) %>% 
  formattable(list(ca_cutoff = normalize_bar("lightgreen", 0.2))) 
```

## Income Cutoffs for Parents in 2010–2018

```{r pa_cutoffs}
melig::parents %>% 
  mutate(
    year = if_else(year == "2009" & month == "December", "2010", year),
    year = year %>% as.integer()
  ) %>% 
  filter(year %in% 2010:2018) %>% 
  select(-month, -state, -fips) %>% 
  tidyr::spread(year, cutoff) %>% 
  rename(
    pa_2010 = `2010`,
    pa_2011 = `2011`,
    pa_2012 = `2012`,
    pa_2013 = `2013`,
    pa_2014 = `2014`,
    pa_2015 = `2015`,
    pa_2016 = `2016`,
    pa_2017 = `2017`,
    pa_2018 = `2018`
  ) %>% 
  formattable(
    list(
     pa_2010 = color_tile("white", "orange"),
     pa_2011 = color_tile("white", "orange"),
     pa_2012 = color_tile("white", "orange"),
     pa_2013 = color_tile("white", "orange"),
     pa_2014 = color_tile("white", "orange"),
     pa_2015 = color_tile("white", "orange"),
     pa_2016 = color_tile("white", "orange"),
     pa_2017 = color_tile("white", "orange"),
     pa_2018 = color_tile("white", "orange")
    )
  ) 
```

## Income Cutoffs for Childless Adults in 2011–2018

```{r ca_cutoffs}
melig::childless_adults %>% 
  mutate(year = as.integer(year)) %>% 
  filter(year %in% 2011:2018) %>% 
  select(-month, -state, -fips) %>% 
  tidyr::spread(year, cutoff) %>% 
  rename(
    ca_2011 = `2011`,
    ca_2012 = `2012`,
    ca_2013 = `2013`,
    ca_2014 = `2014`,
    ca_2015 = `2015`,
    ca_2016 = `2016`,
    ca_2017 = `2017`,
    ca_2018 = `2018`
  ) %>% 
  formattable(
    list(
      ca_2011 = color_tile("white", "orange"),
      ca_2012 = color_tile("white", "orange"),
      ca_2013 = color_tile("white", "orange"),
      ca_2014 = color_tile("white", "orange"),
      ca_2015 = color_tile("white", "orange"),
      ca_2016 = color_tile("white", "orange"),
      ca_2017 = color_tile("white", "orange"),
      ca_2018 = color_tile("white", "orange")
    )
  ) 
```

## Income Cutoffs for Children

_Note_: The children dataset, `melig::children`, has 

- a `type` variable that separates Medicaid and CHIP programs, and 
- an `agegrp` variable that indicates specific age groups.

Because the age information for the CHIP programs (`type == CHIP`) is not available from the source data, I assign a value of "0-18" to the corresponding `age` variable, though some states' CHIP programs don't cover all children under age 19 in some years.

```{r, message=FALSE, warning=FALSE}
# the data
melig::children

# IL
melig::children %>% 
  filter(usps == "IL") 
```

## Income Cutoffs for Pregnant Women and Parents, 2002—2023

```{r Medicaid_Eligibility_Cutoffs_2002-2023, message=FALSE, warning=FALSE, echo=FALSE}
bind_rows(
  melig::parents |> 
  mutate(year = lubridate::ym(paste0(year, month))) |> 
  filter(state == "United States") |> 
  mutate(elig_grp = "Parents"),
  
  melig::pregnant_women |> 
  mutate(year = lubridate::ym(paste0(year, month))) |> 
  filter(state == "United States") |> 
  mutate(elig_grp = "Pregnant Beneficiaries")
) |> 
  mutate(cutoff = cutoff / 100) |> 
  ggplot(aes(x = year, y = cutoff, colour = elig_grp, fill = elig_grp)) +
  geom_point() +
  geom_path() +
  hrbrthemes::theme_ipsum(grid = "XY", plot_title_size = 12) +
  hrbrthemes::scale_y_percent(limits = c(0, 2.5)) +
  theme(legend.position = "top",
        legend.title = element_blank(),
        aspect.ratio = 0.618) +
  labs(x = NULL, y = "Federal Poverl Level",
       title = "Median State-Level Medicaid Income Eligibility Cutoffs, 2002–2023")
```

## Save as Other Data Formats

```r
# save as Stata format
haven::write_dta(melig::parents, "pa0218.dta")

# or
rio::export(melig::parents, "pa0218.dta")
```

Or download the `*.rda` file and try the [rioweb](https://lbraglia.shinyapps.io/rioweb) made by @[lbraglia](https://github.com/lbraglia).

