---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

# Source: https://stackoverflow.com/questions/34983822/how-to-have-r-formattable-rendered-to-pdf-output-and-how-to-have-percents-in-the
export_formattable <- function(f, file, width = "100%", height = NULL, 
                               background = "white", delay = 0.2)
{
  w <- as.htmlwidget(f, width = width, height = height)
  path <- htmltools::html_print(w, background = background, viewer = NULL)
  url <- paste0("file:///", gsub("\\\\", "/", normalizePath(path)))
  webshot::webshot(url,
          file = file,
          selector = ".formattable_widget",
          delay = delay)
}
```

# About `melig`

The R package `melig` contains information on **M**edicaid income **elig**ibility limits, collected by [the Kaiser Family Foundation](http://kff.org/data-collection/trends-in-medicaid-income-eligibility-limits/). It includes the following datasets:

- `children.rda`: 
    - [Medicaid Income Eligibility Limits for Infants Ages 0–1, 2000–2023](https://www.kff.org/medicaid/state-indicator/medicaid-income-eligibility-limits-for-infants-ages-0-1/)
    - [Medicaid Income Eligibility Limits for Children Ages 1–5, 2000–2023](https://www.kff.org/medicaid/state-indicator/medicaid-income-eligibility-limits-for-children-ages-1-5/)
    - [Medicaid Income Eligibility Limits for Children Ages 6–18, 2000–2023](https://www.kff.org/medicaid/state-indicator/medicaid-income-eligibility-limits-for-children-ages-6-18/)
    - [Separate Children’s Health Insurance Program (CHIP) Income Eligibility Limits for Children, 2000–2023](https://www.kff.org/medicaid/state-indicator/separate-childrens-health-insurance-program-chip-income-eligibility-limits-for-children/)
    - [Medicaid/CHIP Upper Income Eligibility Limits for Children, 2000–2023](https://www.kff.org/medicaid/state-indicator/medicaidchip-upper-income-eligibility-limits-for-children/)
    
- `prenant_wom.rda`: [Medicaid and CHIP Income Eligibility Limits for Pregnant Women, 2003–2023](https://www.kff.org/medicaid/state-indicator/medicaid-and-chip-income-eligibility-limits-for-pregnant-women/)
- `parents.rda`: [Medicaid Income Eligibility Limits for Parents, 2002–2023](https://www.kff.org/medicaid/state-indicator/medicaid-income-eligibility-limits-for-parents/)
- `childless_adults.rda`:  [Medicaid Income Eligibility Limits for Other Non-Disabled Adults, 2011–2023](https://www.kff.org/medicaid/state-indicator/medicaid-income-eligibility-limits-for-other-non-disabled-adults/)

The datasets are stored in long format.

The [`/data-raw/`](https://github.com/jjchern/melig/tree/master/data-raw) folder contains raw data and R scripts.

# Installing the Package

You can install `melig` from Github with:

```R
# install.packages("remotes")

# Install the development version
remotes::install_github("jjchern/melig")

# Install the released version
remotes::install_github("jjchern/melig@v0.1.0")

# To uninstall the package, use:
# remove.packages("melig")
```

# Useful Links

- [Medicaid Income Eligibility Limits (Most Recent Data)](http://kff.org/state-category/medicaid-chip/medicaidchip-eligibility-limits/)
- [Annual Updates on Eligibility Rules, Enrollment and Renewal Procedures, and Cost-Sharing Practices in Medicaid and CHIP](http://kff.org/medicaid/report/annual-updates-on-eligibility-rules-enrollment-and/)

# Usage

## Long and Wide Formats

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(formattable)

# List all available datasets
data(package = "melig") %>% 
    pluck("results") %>% 
    as_tibble() %>% 
    select(Package, Item)

# The datasets are in long format
melig::parents

# But you can always convert them back to the original wide format
melig::parents %>% 
  unite_("monyear", c("month", "year"), sep = " ") %>% 
  spread(monyear, cutoff)
```

## Save as Other Data Formats

```r
# save as Stata format
haven::write_dta(melig::parents, "pa0218.dta")

# or
rio::export(melig::parents, "pa0218.dta")
```

Or download the `*.rda` file and try the [rioweb](https://lbraglia.shinyapps.io/rioweb) made by @[lbraglia](https://github.com/lbraglia).

## Explore the Vignette

For a more detailed exploration of the Medicaid income eligibility cutoffs data, check out the [vignette](https://jjchern.github.io/melig/articles/medicaid-income-eligibility-cutoffs-overview.html). To access the vignette within R, after installing the package, you can view the vignette using the following command in your R console:

```r
vignette("medicaid-income-eligibility-cutoffs-overview")
```
