---
title: "Overview of Medicaid Income Eligibility Cutoffs Across States and Eligibility Groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of Medicaid Income Eligibility Cutoffs Across States and Eligibility Groups}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE, 
  echo=FALSE
)
```

```{r setup, message=FALSE, echo=TRUE}
library(tidyverse)
library(formattable)
library(melig)
```

## Income Cutoffs for Pregnant Women and Parents, 2002—2023

```{r Medicaid_Eligibility_Cutoffs_2002-2023}
bind_rows(
  melig::parents |> 
  mutate(year = lubridate::ym(paste0(year, month))) |> 
  filter(state == "United States") |> 
  mutate(elig_grp = "Parents"),
  
  melig::pregnant_women |> 
  mutate(year = lubridate::ym(paste0(year, month))) |> 
  filter(state == "United States") |> 
  mutate(elig_grp = "Pregnant Beneficiaries")
) |> 
  mutate(cutoff = cutoff / 100) |> 
  ggplot(aes(x = year, y = cutoff, colour = elig_grp, fill = elig_grp)) +
  geom_point() +
  geom_path() +
  hrbrthemes::theme_ipsum(grid = "XY", plot_title_size = 12) +
  hrbrthemes::scale_y_percent(limits = c(0, 2.5)) +
  theme(legend.position = "top",
        legend.title = element_blank(),
        aspect.ratio = 0.618) +
  labs(x = NULL, y = "Federal Poverl Level",
       title = "Median State-Level Medicaid Income Eligibility Cutoffs, 2002–2023")
```

## States that Have Income Eligibility Cutoffs Greater Than 100 Federal Poverty Guidelines in 2014

```{r}
# For parents
melig::parents %>% 
  filter(year == 2014) %>% 
  filter(cutoff >= 100) %>% 
  select(state, fips, usps, pa_cutoff = cutoff) %>% 
  formattable(list(pa_cutoff = normalize_bar("lightgreen", 0.2)))

# For childless adults
melig::childless_adults %>% 
  filter(year == 2014) %>% 
  filter(cutoff >= 100) %>% 
  select(state, fips, usps, ca_cutoff = cutoff) %>% 
  formattable(list(ca_cutoff = normalize_bar("lightgreen", 0.2))) 
```

## Income Eligibility Cutoffs for Parents in 2010–2018

```{r pa_cutoffs}
melig::parents %>% 
  mutate(
    year = if_else(year == 2009L & month == "December", 2010L, year),
    year = year %>% as.integer()
  ) %>% 
  filter(year %in% 2010:2018) %>% 
  select(-month, -state, -fips) %>% 
  tidyr::spread(year, cutoff) %>% 
  rename(
    pa_2010 = `2010`,
    pa_2011 = `2011`,
    pa_2012 = `2012`,
    pa_2013 = `2013`,
    pa_2014 = `2014`,
    pa_2015 = `2015`,
    pa_2016 = `2016`,
    pa_2017 = `2017`,
    pa_2018 = `2018`
  ) %>% 
  formattable(
    list(
     pa_2010 = color_tile("white", "orange"),
     pa_2011 = color_tile("white", "orange"),
     pa_2012 = color_tile("white", "orange"),
     pa_2013 = color_tile("white", "orange"),
     pa_2014 = color_tile("white", "orange"),
     pa_2015 = color_tile("white", "orange"),
     pa_2016 = color_tile("white", "orange"),
     pa_2017 = color_tile("white", "orange"),
     pa_2018 = color_tile("white", "orange")
    )
  ) 
```

## Income Eligibility Cutoffs for Childless Adults in 2011–2018

```{r ca_cutoffs}
melig::childless_adults %>% 
  mutate(year = as.integer(year)) %>% 
  filter(year %in% 2011:2018) %>% 
  select(-month, -state, -fips) %>% 
  tidyr::spread(year, cutoff) %>% 
  rename(
    ca_2011 = `2011`,
    ca_2012 = `2012`,
    ca_2013 = `2013`,
    ca_2014 = `2014`,
    ca_2015 = `2015`,
    ca_2016 = `2016`,
    ca_2017 = `2017`,
    ca_2018 = `2018`
  ) %>% 
  formattable(
    list(
      ca_2011 = color_tile("white", "orange"),
      ca_2012 = color_tile("white", "orange"),
      ca_2013 = color_tile("white", "orange"),
      ca_2014 = color_tile("white", "orange"),
      ca_2015 = color_tile("white", "orange"),
      ca_2016 = color_tile("white", "orange"),
      ca_2017 = color_tile("white", "orange"),
      ca_2018 = color_tile("white", "orange")
    )
  ) 
```

## Income Eligibility Cutoffs for Children

_Note_: The `melig::children` dataset includes two key variables:

- `type`: Distinguishes between Medicaid and CHIP (Children's Health Insurance Program) programs.
- `agegrp`: Indicates specific age groups within these programs.

The age data for CHIP programs (`type` == CHIP) is not explicitly provided in the source data. Therefore, in this dataset, a general age range of "0-18" is assigned to the age variable for CHIP. However, certain states' CHIP programs may not cover all children under age 19 in some years.

```{r, message=FALSE, warning=FALSE}

melig::children

# IL
melig::children %>% 
  filter(usps == "IL") 
```
